---
title: "Biostat 203B Homework 5"
subtitle: Due Mar 22 @ 11:59PM
author: "Chengwu Duan and 606332825"
format:
  html:
    theme: cosmo
    embed-resources: true
    number-sections: false
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
---

## Predicting ICU duration

Using the ICU cohort `mimiciv_icu_cohort.rds` you built in Homework 4, develop at least three machine learning approaches (logistic regression with enet regularization, random forest, boosting, SVM, MLP, etc) plus a model stacking approach for predicting whether a patient's ICU stay will be longer than 2 days. You should use the `los_long` variable as the outcome. Your algorithms can use patient demographic information (gender, age at ICU `intime`, marital status, race), ICU admission information (first care unit), the last lab measurements before the ICU stay, and first vital measurements during ICU stay as features. You are welcome to use any feature engineering techniques you think are appropriate; but make sure to not use features that are not available at an ICU stay's `intime`. For instance, `last_careunit` cannot be used in your algorithms. 

### 1. Data preprocessing and feature engineering.

**Answer: ** Done in hw5Preprocess.qmd, will be using the result rds here.

```{r}
#| message: false
#| warning: false
#| error: false

# Load libraries
library(GGally)
library(gtsummary)
library(ranger)
library(keras)
library(stacks)
library(tidyverse)
library(tidymodels)
library(vip)
library(ggplot2)

# load data
mimiciv_icu_cohort <- read_rds("./mimiciv_icu_cohort_postprocessed.rds")
```

### 2. Partition data into 50% training set and 50% test set. Stratify partitioning according to `los_long`. For grading purpose, sort the data by `subject_id`, `hadm_id`, and `stay_id` and use the seed `203` for the initial data split. Below is the sample code.

```{r}
set.seed(203)

# sort
mimiciv_icu_cohort <- mimiciv_icu_cohort %>%
  arrange(subject_id, hadm_id, stay_id)

data_split <- initial_split(
  mimiciv_icu_cohort, 
  # stratify by los_long
  strata = "los_long", 
  prop = 0.5
  )

mimic_train <- training(data_split)
dim(mimic_train)

mimic_test <- testing(data_split)
dim(mimic_test)

# checking for missing values
summary(mimiciv_icu_cohort)

mimiciv_icu_cohort %>% tbl_summary()
```

### 3. Train and tune the models using the training set.

#### Multi-layer perceptron (MLP) - Technical difficulty with Python reticulate DOES NOT WORK

```{r, eval = F}
library(doParallel)

# eval = F to only have to run model once after tuning process
mlp_recipe <- 
  recipe(
    los_long ~ ., 
    data = mimic_train
  ) %>%
  # mean imputation for NAs in numeric predictors
  step_impute_mean(all_numeric_predictors()) %>%
  # mode imputation for NAs in categorical predictors
  step_impute_mode(all_nominal_predictors()) %>%
  # create traditional dummy variables
  step_dummy(all_nominal_predictors()) %>%
  # zero-variance filter
  step_zv(all_numeric_predictors()) %>%
  # center and scale numeric data
  step_normalize(all_numeric_predictors())
mlp_recipe

mlp_mod <- 
  mlp(
    mode = "classification",
    hidden_units = tune(),
    dropout = tune(),
    epochs = 10,
  ) %>% 
  set_engine("keras", verbose = 0)
mlp_mod


mlp_wf <- workflow() %>%
  add_recipe(mlp_recipe) %>%
  add_model(mlp_mod)
mlp_wf

write_rds(mlp_wf, "mlp_wf.rds")

param_grid <- grid_regular(
  hidden_units(range = c(1, 3)),
  dropout(range = c(0, 0.6)),
  levels = 3
  )
param_grid

set.seed(203)

folds <- vfold_cv(mimic_train, v = 3)
folds

# saving 1 cpu for other tasks
registerDoParallel(cores = detectCores()-2)

(mlp_fit <- mlp_wf %>%
  tune_grid(
    resamples = folds,
    grid = param_grid,
    metrics = metric_set(roc_auc, accuracy)
    )) %>%
  system.time()

# stop the registered parallel backend
stopImplicitCluster()

mlp_fit

write_rds(mlp_fit, "mlp_fit.rds")
```

```{r}
mlp_fit <- read_rds("mlp_fit.rds")
mlp_wf <- read_rds("mlp_wf.rds")

mlp_fit %>%
  show_best(metric = "roc_auc")
mlp_fit

best_mlp <- mlp_fit %>%
  select_best(metric = "roc_auc")
best_mlp

mlp_fit %>%
  collect_metrics() %>%
  print(width = Inf) %>%
  filter(.metric == "roc_auc") %>%
  ggplot(mapping = aes(x = dropout, y = mean, 
                       color = factor(hidden_units))) +
    geom_point() +
    geom_line() +
    labs(x = "Dropout Rate", y = "CV AUC") +
    scale_x_log10()

# Final workflow
final_wf <- mlp_wf %>%
  finalize_workflow(best_mlp)
final_wf

# Fit the whole training set, then predict the test cases
final_fit <- 
  final_wf %>%
  last_fit(data_split)
final_fit

# Test metrics
final_fit %>% 
  collect_metrics()
```

```{r, eval = F}
# saving results for stacked model
mlp_res <- 
  tune_grid(
    object = mlp_wf, 
    resamples = folds, 
    grid = param_grid,
    control = control_stack_grid()
  )
write_rds(mlp_res, "mlp_res.rds")
```

### 4. Compare model classification performance on the test set. Report both the area under ROC curve and accuracy for each machine learning algorithm and the model stacking. Interpret the results. What are the most important features in predicting long ICU stays? How do the models compare in terms of performance and interpretability?

#### 4.1 The importance of features in predicting long ICU stays

```{r}
library(vip)

#library(rpart.plot)
final_tree <- extract_workflow(final_fit)
final_tree

final_tree %>% 
  extract_fit_parsnip() %>% 
  vip()
```

**Answer: ** first_careunit is the most important feature in predicting long ICU stays, followed by heart rate of the patient then the admission location.

#### 4.2 Interpretation of the results

**Answer: ** The multi-layer perceptron model has an test AUC of 0.607 and an accuracy of 0.579. The most important features in predicting long ICU stays are first_careunit, heart rate, and admission location. The model is interpretable and has a decent performance.