---
title: "Biostat 203B Homework 5"
subtitle: Due Mar 22 @ 11:59PM
author: "Chengwu Duan and 606332825"
format:
  html:
    theme: cosmo
    embed-resources: true
    number-sections: false
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
---

## Predicting ICU duration

Using the ICU cohort `mimiciv_icu_cohort.rds` you built in Homework 4, develop at least three machine learning approaches (logistic regression with enet regularization, random forest, boosting, SVM, MLP, etc) plus a model stacking approach for predicting whether a patient's ICU stay will be longer than 2 days. You should use the `los_long` variable as the outcome. Your algorithms can use patient demographic information (gender, age at ICU `intime`, marital status, race), ICU admission information (first care unit), the last lab measurements before the ICU stay, and first vital measurements during ICU stay as features. You are welcome to use any feature engineering techniques you think are appropriate; but make sure to not use features that are not available at an ICU stay's `intime`. For instance, `last_careunit` cannot be used in your algorithms. 

### 1. 

Data preprocessing and feature engineering.

**Answer: ** Done in hw5Preprocess.qmd, the output mimiciv_icu_cohort_postprocessed.rds will be used here as source rds.

```{r}
#| message: false
#| warning: false
#| error: false

# Load libraries
library(GGally)
library(gtsummary)
library(ranger)
library(keras)
library(stacks)
library(tidyverse)
library(tidymodels)
library(vip)
library(ggplot2)

# load data
mimiciv_icu_cohort <- read_rds("./mimiciv_icu_cohort_postprocessed.rds")
```

### 2. 

Partition data into 50% training set and 50% test set. Stratify partitioning according to `los_long`. For grading purpose, sort the data by `subject_id`, `hadm_id`, and `stay_id` and use the seed `203` for the initial data split. Below is the sample code.

```{r}
set.seed(203)

# sort
mimiciv_icu_cohort <- mimiciv_icu_cohort %>%
  arrange(subject_id, hadm_id, stay_id)

data_split <- initial_split(
  mimiciv_icu_cohort, 
  # stratify by los_long
  strata = "los_long", 
  prop = 0.5
  )

mimic_train <- training(data_split)
dim(mimic_train)

mimic_test <- testing(data_split)
dim(mimic_test)

# checking for missing values
summary(mimiciv_icu_cohort)

mimiciv_icu_cohort %>% tbl_summary()
```

### 3. 

Train and tune the models using the training set.

#### Stacked model

```{r}
stack_recipe <- 
  recipe(
    los_long ~ ., 
    data = mimic_train
  ) %>%
  # mean imputation for NAs in numeric predictors
  step_impute_mean(all_numeric_predictors()) %>%
  # mode imputation for NAs in categorical predictors
  step_impute_mode(all_nominal_predictors()) %>%
  # create traditional dummy variables
  step_dummy(all_nominal_predictors()) %>%
  # zero-variance filter
  step_zv(all_numeric_predictors()) %>%
  # center and scale numeric data
  step_normalize(all_numeric_predictors())
stack_recipe
```

```{r, eval = FALSE}
library(doParallel)

logit_res <- read_rds("logit_res.rds")
gb_res <- read_rds("gb_res.rds")
mlp_res <- read_rds("mlp_res.rds")

# Register doParallel as the parallel backend as Random Forest is computationally expensive
# saving 2 cpu for other tasks
registerDoParallel(cores = detectCores()-2)

stack_model_st <- 
  # initialize the stack
  stacks() %>%
  # add candidate members
  add_candidates(logit_res) %>%
  add_candidates(gb_res) %>%
  add_candidates(mlp_res) %>%
  # determine how to combine their predictions
  blend_predictions(
    penalty = 10^(-7:2),
    metrics = c("roc_auc")
    ) %>%
  # fit the candidates with nonzero stacking coefficients
  fit_members()

# stop the registered parallel backend
stopImplicitCluster()
write_rds(stack_model_st, "stack_model_st.rds")
```

```{r}
library(yardstick)
library(tidyverse)
library(ggplot2)
library(ggfortify)

stack_model_st = read_rds("stack_model_st.rds")

autoplot(stack_model_st)
autoplot(stack_model_st, type = "members")
autoplot(stack_model_st, type = "weights")
collect_parameters(stack_model_st, "logit_res")
collect_parameters(stack_model_st, "gb_res")
collect_parameters(stack_model_st, "mlp_res")
```

```{r}
library(yardstick)

stack_pred <- mimic_test %>%
  bind_cols(predict(stack_model_st, ., type = "prob")) %>%
  print(width = Inf)

roc_auc(
  stack_pred,
  truth = los_long,
  contains(".pred_FALSE")
  )
```

```{r, eval = F}
# Register doParallel as the parallel backend as Random Forest is computationally expensive
# saving 2 cpu for other tasks
registerDoParallel(cores = detectCores()-2)

stack_pred <-
  mimic_test %>%
  select(los_long) %>%
  bind_cols(
    predict(
      stack_model_st,
      mimic_test,
      type = "class",
      members = TRUE
      )
    ) %>%
  print(width = Inf)

# stop the registered parallel backend
stopImplicitCluster()
write_rds(stack_pred, "stack_model_pred.rds")
```

```{r}
library(yardstick)
library(tidyverse)
stack_pred = read_rds("stack_model_pred.rds")

stack_pred

map(
  colnames(stack_pred),
  ~mean(stack_pred$los_long == pull(stack_pred, .x))
  ) %>%
  set_names(colnames(stack_pred)) %>%
  as_tibble() %>%
  pivot_longer(c(everything(), -los_long))
```

### 4. 

Compare model classification performance on the test set. Report both the area under ROC curve and accuracy for each machine learning algorithm and the model stacking. Interpret the results. What are the most important features in predicting long ICU stays? How do the models compare in terms of performance and interpretability?

**Answer: ** The test AUC for the stacked model is 0.659. Relatively compared to other individual models, the stacked model has the highest AUC.
